\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-07-29 paper/paper.tex
Updates.

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are expensive - retailing at approximately 2X the price of a standard DIMM [find
  ref]. This means that they will likely be available in small quantities. We
compensate for this by periodically draining the NVDIMM log into a log placed on
a standard SSD on the system.

@@

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. While we designed Cyclone to leverage
new non-volatile memory technologies that might be available in the
future~\cite{3dxpoint}, current non-volatile memory options such are NVDIMMs are
more expensive that DRAM. Therefore, for Cyclone to be economical we cannot
assume that a sufficient amount of NVM is available to hold the entire log. The
log entries can only be garbage collected when the in-memory component of the
key value store is checkpointed to secondary storage. The growing size of
available DRAM means that also provisioning sufficient NVDIMMs to hold the
corresponding log of updates is not feasible given that NVDIMMs cost at least
twice as much as volatile DRAM. We therefore resort to using a second larger log
level placed on a flash SSD. Entries are drained from the NVM log to this flash
log in conveniently large units that are a multiple of 4KB - the optimum IO unit
for flash SSDs. This ensures we get the best possible throughput from a flash
device without paying the price of synchronous IO for small key-value pairs. At
the same time, the amount of non-volatile memory required is small enough to not
add undue cost of the server.

>>>>

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are significantly more expensive than standard DIMMs. This means that they will
likely be available in small quantities. We compensate for this by periodically
draining the NVDIMM log into a log placed on a standard SSD on the system.

@@

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. While we designed Cyclone to leverage
new non-volatile memory technologies that might be available in the
future~\cite{3dxpoint}, current non-volatile memory options such are NVDIMMs are
more expensive that DRAM. Therefore, for Cyclone to be economical we cannot
assume that a sufficient amount of NVM is available to hold the entire log. The
log entries can only be garbage collected when the in-memory component of the
key value store is checkpointed to secondary storage. The growing size of
available DRAM means that also provisioning sufficient NVDIMMs to hold the
corresponding log of updates is not feasible given that NVDIMMs are
significantly more expensive than standard volatile DIMMs.  To illustrate this,
at the time of writing this paper, an 8 GB DDR4 NVDIMM was retailing at \$612,
while a plain 8 GB DD4 DIMM was retailing at \$178.  We therefore resort to
using a second larger log level placed on a flash SSD. Entries are drained from
the NVM log to this flash log in conveniently large units that are a multiple of
4KB - the optimum IO unit for flash SSDs. This ensures we get the best possible
throughput from a flash device without paying the price of synchronous IO for
small key-value pairs. At the same time, the amount of non-volatile memory
required is small enough to not add undue cost of the server.
